---
title: "Final Project"
author: "Nina Cohen"
date: "5/17/2018"
output: github_document
---

1. Introduction

I chose to work with the Global Terrorism Database because terrorism and terrorist attacks are a very relevant threat and topic in the news and socially. Growing up with my father working for the Department of Homeland Security and counterterrorism, I have always been fascinated with the the topic and learning more. Terrorism is such a major problem globally right now and using data science to possibly predict or support new policies that help protect against future attacks or situations.

In this script I will cover:
  1) Data Curation
  2) Parsing Data
  3) Management of Data
  4) Exploratory Data Analysis (EDA) 
  5) Hypothesis Testing and Machine Learning 


1.1 Load Data

To load the data set or data of interest, first download the csv file from a website, such as https://www.kaggle.com/datasets and save it to the same folder as the file you will be using it in. Then you want to read the csv file and save it to a variable so you can later access and manipulate the data.

```{r setupdb, include=TRUE,warning=FALSE, message=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
db <- read_csv("globalterrorism.csv")
db[1:10,1:7]
```

2. Data Curation

In this code, there is a lot of repeated information, or unneccesary information for what I plan to do with the information. For example, the eventid number is just a combination of the year, month, day, extended value and some count value of somesort. For future information, having a distinct eventid or reference id for each even will not be useful.

The approxdate gives a range of dates the attack took place over, resolution I assume provides a date where a resolution to the attack was created, though in the description of the data set there is no index that describes what each column represents. In all, the main columns that will be important to us will be the year, month, day, country/country id number, provstate, city, type of attack (we will only be focusing on the primary attack of each individual attack, thus will only get the information from attacktype1 and such) and target type.

Below is the code where I select only these column from the database. I also only select rows where the year is after 2001 so I can use more recent data to calcuate future predictions or information.


```{r dataCuration, warning=FALSE, message=FALSE}

tidy_db <- db %>% 
  filter(iyear >= 2001) %>%
  dplyr::select(iyear,imonth,iday,country,country_txt,region,region_txt,provstate,city,attacktype1,
                attacktype1_txt,targtype1,targtype1_txt) 

tidy_db %>% head(10)
```

3. Parsing Data


Here, I am selecting and graphin the number of attacks in each city per year for my ten cities of choice: DC, NYC, London, Jerusalem, Paris, Baghdad, New Delhi, Grozny, Gaza and Barrancabermeja. I chose these cities becasue the majority of them are major cities in countries where terrorism is a major threat, or as I was scrolling through the database, I constantly saw the name reappear. Here I will get a database that contains the number of attacks for all of the cities in the database, numAttacks and numAttacksPerCity, and for just my specified cities, numAttacksMainCities and numAttacksPerCity2, so I can analyze the data differently later on. Column x represents the number of attacks for that year in that specific city.

```{r dataParsing1}

numAttacks <- tidy_db %>%
  group_by(iyear, country_txt, city) %>%
  dplyr::select(iyear,country_txt,city,attacktype1) %>%
  dplyr::count(attacktype1)

numAttacksMainCities <- numAttacks %>%
  filter(city == "District of Columbia" || city == "New York City" || city == "London"||city == "Jerusalem"
         || city == "Paris"||city == "Baghdad"||city == "New Delhi"||city == "Grozny"
         ||city == "Gaza"||city == "Barrancabermeja") 


numAttacksPerCity <- aggregate(numAttacks$n, by=list(city=numAttacks$city,iyear=numAttacks$iyear), FUN=sum) 
numAttacksPerCityMain <- aggregate(numAttacksMainCities$n,
                      by=list(city=numAttacksMainCities$city,iyear=numAttacksMainCities$iyear), FUN=sum) 

numAttacksPerCity %>% head(10)
numAttacksPerCityMain %>% head(10)
```

Here I will look at the number of attacks for all of the cities globally to see which city had the most attacks overall. Column x represents the number of attacks for that year in that specific city.

```{r dataParsing2}

mostAttacksPerYear2 <- numAttacksPerCity %>% 
  group_by(iyear) %>%
  filter(x == max(x))

mostAttacksPerYear2

```

Now, I wil calculate the number of attacks per country and divide it by type of attack. The numbers, 1 - 9, each represent a different type of attack. The legend for what each number stands for can be seen in the console. 1 is an Assasination, 2 is an Armed Assault, 3 is a Bombing/Explosion, 4 is a Hijacking, 5 is a Hostage Taking (Barricade), 6 is a Hostage Taking (Kidnapping), 7 is a Facility/Infastructure Attack, 8 is an Unarmed Assault and finally, 9 is Unknown.
```{r dataParsing3}

numAttacksCountryType <- tidy_db %>%
  group_by(iyear, country_txt) %>%
  dplyr::select(iyear,country_txt,attacktype1,attacktype1_txt) %>%
  dplyr::count(attacktype1) 

numAttacksCountryType %>% head(10)
    
numAttacksPerCountry <- aggregate(numAttacksCountryType$n,
            by = list(country=numAttacksCountryType$country_txt,year=numAttacksCountryType$iyear), FUN=sum) 

numAttacksPerCountry %>% head(10)


numAttacksCountryType %>%
  ggplot(aes(x=iyear, y=n)) +
    geom_point() +
    labs(title="Number of Attacks a Year Per Country",
         x = "Year",
         y = "Number of Attacks") + facet_grid(.~attacktype1) 
  labs(caption = "1: Assassination, 2: Armed Assault, 3: Bombing/Explosion, 4: Hijacking, 
       5: Hostage Taking (Barricade), 6: Hostage Taking (Kidnapping), 7: Facility/Infastructure Attack,
       8: Unarmed Assault, 9: Unknown")


```
As you can see, looking at all of the countries globally creates a graph that does not give us much information. Thus, I will look at the data for the countries where my specified cities are located.

```{r dataParsing3b}

numAttacksCountryType <- tidy_db %>%
  group_by(iyear, country_txt,city) %>%
  dplyr::select(iyear,country_txt,city,attacktype1,attacktype1_txt) %>%
  filter(country_txt == "United States" || country_txt == "United Kingdom"||country_txt == "Israel"
  || country_txt == "France"||country_txt == "Iraq"||country_txt == "India"||country_txt == "Russia"
  ||country_txt == "West Bank and Gaza Strip"||country_txt == "Colombia") %>%
  dplyr::count(attacktype1) 

numAttacksCountryType %>% head(10)

numAttacksPerCountry <- aggregate(numAttacksCountryType$n,
  by = list(country=numAttacksCountryType$country_txt,iyear=numAttacksCountryType$iyear), FUN=sum) 

numAttacksPerCountry %>% head(10)
```
The first table shows us the number of attacks per country per type of attack. Seeing the data broken down like this can give governments or officials better information about how to protect their citizens against types of attacks if they know which type is the most common. The second table shows the total number of attacks per country. Seeing all the types of terrorist attacks grouped together allows one to interpret which country is at the biggest risk for attacks in the years to come.

4. Management of Data

Having null or unknown rows, as seen in the chart with the max number of attacks per year in a city, is not useful and does not give the user or reader any information. Removing unknowns or NA's will give more information that is accurate and critical to future analysis. Here, we will remove any city that is Unknown and use this table for any analysis that involves solely cities, then use the previous table, numAttacksPerCountry, when doing analysis based on country. 

```{r managementOfData}

cleaned_db <- filter(numAttacksPerCity,numAttacksPerCity$city != "Unknown") 

cleaned_db %>% head(10)

mostAttacksPerYear <- cleaned_db %>% 
  group_by(iyear) %>%
  filter(x == max(x))

mostAttacksPerYear
```
Now, we can see that two of the ten major cities we were looking at are the two cities that had the most attacks each year from 2001 - 2016.

5. Exploratory Data Analysis (EDA) 

First, lets graph the number of attacks since 2001 in the United States to find which states had the most attacks.
```{r EDA1}

USattacks <- filter(tidy_db,tidy_db$country_txt == "United States") %>%
  group_by(provstate, attacktype1,attacktype1_txt) %>%
  dplyr::select(provstate,attacktype1, attacktype1_txt) %>%
  dplyr::count(attacktype1) 

USattacks2 <- aggregate(USattacks$n, by=list(state=USattacks$provstate), FUN=sum) 

USattacks2 <- USattacks2[order(-USattacks2$x),] %>% head(10)
USattacks2

USattacks2 %>%
  ggplot(aes(x=state, y=x, color = state)) +
    geom_point() +
    labs(title="Number of Attacks a Year Per State From 2001 - 2016",
         x = "State",
         y = "Number of Attacks")


```
From this graph and table, we can conclude that since 2001, California and New York have had the most attacks. Knowing this, the Government and police forces should take this into account when training or putting funds within Homeland Security because these are the two areas/states where peopel are at the most risk for being affected by a terrorist attack

Now, we will graph the number of total attacks per year for each of the ten cities we are analyzing. 
```{r EDA2}

numAttacksPerCityMain %>%
  ggplot(aes(x=iyear, y=x, color = city)) +
    geom_point() +
    labs(title="Number of Attacks a Year Per City",
         x = "Year",
         y = "Number of Attacks") 

```
As you can see, Baghdad has a significantly higher number of attacks than all of the other cities. This could be due to the fact that the US invaded Iraq in 2003 and was fighting a war there till 2011.  


Next, we will graph the number of attacks per contry by type of attack for each of the countries where the ten cities we are focusing on is located. 
```{r EDA3}
numAttacksCountryType

numAttacksCountryType %>%
  ggplot(aes(x=iyear, y=n, color = country_txt)) +
    geom_point() +
    labs(title="Number of Attacks a Year Per Country",
         x = "Year",
         y = "Number of Attacks") + facet_grid(attacktype1~.)

```


6. Hypothesis Testing and Machine Learning 

As you can see from the graph that displays the number of attacks per year in the ten cities we are focusing on, overall Baghdad, Jerusalem and Gaza have the most attacks. Using this knowledge, we will predict the number of attacks for the following year.

```{r hypoth1}

newCities <- numAttacksCountryType %>% 
  filter(country_txt == "Iraq" || country_txt == "Israel" || country_txt == "West Bank and Gaza Strip")

lmPlot2 <- lm(n ~ country_txt*iyear, data = newCities)
lmPlot <- lmPlot2 %>% broom::tidy()
lmPlot

increaseIraq <- lmPlot$estimate[4]
increaseIsrael <- lmPlot$estimate[4] + lmPlot$estimate[5]
increaseWBGaza <- lmPlot$estimate[4] + lmPlot$estimate[6]

increaseIraq
increaseIsrael
increaseWBGaza
```
From this table, we learn that the intercept is the attacks per year for Iraq, which means each year the number of attacks will increase by about .06, or increaseIraq. The number of attacks in Israel, increaseIsrael, will increase by .022 attacks per year. Finally, the number of attacks in the West Bank and Gaza will decrease by .0209 attcks per year. 
